---
title: "Template for 'Financial statements: A first look'"
author: Your name
format: html
---

## Setting up WRDS

```{r}
#| include: false
library(dplyr, warn.conflicts = FALSE)
library(DBI)
library(ggplot2)
library(farr)
```

```{r}
Sys.setenv(DATA_DIR = "~/Dropbox/pq_data")
```

```{r}
#| include: false
db <- dbConnect(duckdb::duckdb())
```

## Financial statement data

## Exercises

1. Suppose we didn't have access to Compustat (or an equivalent database) for the analysis above, describe a process you would use to get the data required to make the plot above comparing performance of Microsoft and Apple.

It does not seem easy to get these data from either Google Finance or Yahoo Finance.
We could hand-collect these items from the financial statements of the firms, but not easily before ~1995.
Bloomberg or similar might be another source.
The reality is that it would be very difficult to collect these data without access to a commercial database (Compustat data are supplied in various forms to private industry; FactSet is an alternative).

2. In the following code, how do `funda_mod` and `funda_mod_alt` differ?
(For example, where are the data for each table?)
What does the statement `collect(n = 100)` at the end of this code do?

```{r}
#| include: false
funda <- load_parquet(db, "funda", "comp")

funda_mod <-
  funda |>
  filter(indfmt == "INDL", datafmt == "STD",
         consol == "C", popsrc == "D") |>
  filter(fyear >= 1980)

funda_mod_alt <-
  funda_mod |>
  collect(n = 100)
```

3. The table `comp.company` has data on SIC (Standard Industrial Classification) codes in the field `sic`.
In words, what is the `case_when` function doing in the following code?
Why do we end up with just two rows?

```{r}
company <- load_parquet(db, "company", "comp")

sample <-
  company |>
  select(gvkey, sic) |>
  mutate(co_name = case_when(gvkey == "001690" ~ "Apple",
                             gvkey == "012141" ~ "Microsoft",
         TRUE ~ NA_character_)) |>
  filter(!is.na(co_name))
```

4. What does the data frame `another_sample` represent?
What happens if we change the `inner_join` statement below to simply `inner_join(sample)`.
What happens if we omit change it to `inner_join(sample, by = "sic")` (i.e., omit the `suffix = c("", "_other")` portion)?
Why do you think we want `filter(gvkey != gvkey_other)`?

```{r}
another_sample <-
  company |>
  select(gvkey, sic) |>
  inner_join(sample, by = "sic", suffix = c("", "_other")) |>
  filter(gvkey != gvkey_other) |>
  mutate(group = paste(co_name, "peer")) |>
  select(gvkey, group)
```

5. What is the following code doing?

```{r}
total_sample <-
  sample |>
  rename(group = co_name) |>
  select(gvkey, group) |>
  union_all(another_sample)
```

6. Suppose that we are interested in how firms' level of R&D activity. 
One measure of R&D activity is *R&D Intensity*, which can be defined as "R&D expenses" (Compustat item `xrd`) scaled by "Total Assets" (Compustat item `at`). 
In `xrd_at`, what's the difference between `rd_intensity` and `rd_intensity_alt`?
Does `filter(at > 0)` seem like a reasonable filter?
What happens if we omit it?

```{r}
xrd_at <-
  funda_mod |>
  select(gvkey, datadate, fyear, conm, xrd, at) |>
  filter(at > 0) |>
  mutate(rd_intensity = xrd/at,
         xrd_alt = coalesce(xrd, 0),
         rd_intensity_alt = xrd_alt/at) |>
  inner_join(total_sample, by = "gvkey") 
```

7. Looking at a sample of rows from `xrd_at_sum`, it appears that the three R&D intensity measures are always identical for Apple and Microsoft, but generally different for their peer groups.
What explains these differences?
Can you say that one measure is "correct"? 
Or would you say "it depends"?

```{r}
xrd_at_sum <-
  xrd_at |>
  group_by(group, fyear) |>
  summarize(total_at = sum(at, na.rm = TRUE),
            total_rd = sum(xrd, na.rm = TRUE),
            rd_intensity1 = mean(xrd/at, na.rm = TRUE),
            rd_intensity2 = mean(xrd_alt/at, na.rm = TRUE),
            .groups = "drop") |>
  mutate(rd_intensity3 = if_else(total_at > 0, total_rd/total_at, NA_real_)) |>
  collect()
  
xrd_at_sum |>
  select(-total_at, -total_rd) |>
  arrange(desc(fyear), group)
```

8. Write code to produce the following plot (see [book](https://iangow.github.io/far_book/fin-state.html#exercises-8) for the plot).
Also produce plots that use `rd_intensity1` and `rd_intensity2` as measures of R&D intensity.
Do the plots help you think about which of the three measures makes most sense?

```{r}
#| include: false
dbDisconnect(db, shutdown=TRUE)
```