---
title: "Exercise template for 'Natural experiments'"
author: Your name
format: html
bibliography: book.bib
---

```{r}
#| message: false
library(farr)
library(ggplot2)
library(dplyr, warn.conflicts = FALSE)
```

```{r}
michels_2017 |> 
  mutate(days_to_filing = as.integer(date_filed - eventdate)) |>
  ggplot(aes(x = days_to_filing, fill = recognize)) + 
  geom_histogram(binwidth = 1) 
```

```{r}
michels_2017 |> 
  mutate(days_to_period_end = as.integer(next_period_end - eventdate)) |>
  ggplot(aes(x = days_to_period_end, fill = recognize)) + 
  geom_histogram(binwidth = 1)
```

```{r}
michels_2017 |> 
  count(eventdate) |>
  arrange(desc(n)) |>
  top_n(5)
```

```{r}
#| include: false
top5 <-
  michels_2017 |> 
  count(eventdate) |>
  arrange(desc(n)) |>
  top_n(5) |>
  pull() |>
  sum()
```

## Discussion questions

1. One assumption in @Michels:2017uc is that whether a natural disaster occurs before or after the balance sheet date of the next filing is random.
Do the inherent properties of natural disasters ensure that they are random?
Why?
If not, how would you evaluate the randomness of natural disasters in the sample of @Michels:2017uc?
Do the analyses above help this evaluation?
2. Describe what you imagine to be the process from the occurrence of a natural disaster to reporting on that event in the subsequent filing? Do you think this process differs for recognized versus disclosed events?
3. From the analysis, above it appears that five natural disasters account for `r top5` observations. 
A simple Google search for each date and the word "disaster" reveals that these events are Hurricane Katrina (`2005-08-29`), Hurricane Ike (`2008-09-13`), Hurricane Ivan (`2004-09-16`), Hurricane Charley (`2004-08-13`), and Hurricane Wilma (`2005-10-24`).
Is it problematic that a small number of disasters accounts for a significant portion of the sample?
3. Where does @Michels:2017uc get data on natural disasters from?
Is there anything that is problematic about this data source?
Would it be possible to use another approach to data collection?
What challenges would that approach face?
4. A recurring question in accounting is whether it matters whether information is disclosed or recognized? 
One view is that, if markets are efficient, it should not matter where the information is disclosed, so recognition should not matter relative to disclosure.
What assumptions underlie this view?
Are there any reasons to believe that they do or do not hold in the setting of @Michels:2017uc?
What are the implications for the ability of @Michels:2017uc to deliver clean causal inferences?
Nevertheless, even in this relatively straightforward setting, @Michels:2017uc recognizes the possibility of different materiality criteria for disclosed and recognized events, which could affect the relation been underlying events and observed disclosures. 
@Michels:2017uc takes care to address this concern.
5. What causal inferences does @Michels:2017uc draw?
What (if any) issues do you see with regard to these?
6. Choose a paper that you have seen recently that uses empirical analysis of non-experimental data.
(If you cannot choose such a paper, you could use @Hopkins:2018tg, which was a seminar paper recently at the University of Melbourne.)
Looking at the abstract of the paper can you determine whether this paper seeks to draw causal inferences?
7. Choose what you think the authors regard to be the most important causal inference they draw (or would like to draw) in your chosen paper. 
Which table or tables provide the relevant empirical analyses?
Sketch a rough causal diagram for this causal inference using either discussion in the paper or your own background knowledge to identify important variables.
How credible do you find the reported causal inferences to be in light of your causal diagram?
