---
title: "Exercise template for 'Panel data'"
author: Your name
format: html
bibliography: book.bib
---

```{r}
#| warning: false
#| message: false
#| include: false
library(dplyr, warn.conflicts = FALSE)
library(farr)
library(DBI)
library(lubridate)
library(fixest)
library(modelsummary)
library(ggplot2)
```

```{r}
#| echo: false
db <- dbConnect(RPostgres::Postgres(), bigint = "integer")

seg_customer <- tbl(db, Id(schema = "compseg", table = "seg_customer"))
names_seg <- tbl(db, Id(schema = "compseg", table = "names_seg"))
funda <- tbl(db, Id(schema = "comp", table = "funda"))
```

```{r}
#| echo: false
undisclosed_names <-
  undisclosed_names |>
  distinct()

customers <-
    seg_customer |>
    filter(ctype == "COMPANY") |>
    collect()

disclosure_raw <-
    customers |>
    filter(between(datadate, as.Date("1994-01-01"), as.Date("2010-12-31"))) |>
    left_join(undisclosed_names, by = "cnms") |>
    mutate(disclosed = coalesce(disclosed, TRUE)) |>
    select(gvkey, datadate, cnms, salecs, disclosed) 
```

```{r}
#| echo: false
sales <-
  funda |>
  filter(indfmt == "INDL", datafmt == "STD",
         consol == "C", popsrc == "D") |>
  select(gvkey, datadate, sale) |>
  collect()
```

```{r}
#| echo: false
prin_cust_df <-
  disclosure_raw |> 
  inner_join(sales, by = c("gvkey", "datadate")) |>
  group_by(gvkey, datadate) |> 
  filter(!is.na(salecs), sale > 0) |>
  summarize(prin_cust = max(salecs/sale, na.rm = TRUE), 
            .groups = 'drop') |>
  mutate(has_prin_cust = prin_cust >= 0.1)
```

```{r}
#| echo: false
disclosure <-
    disclosure_raw |> 
    inner_join(sales, by = c("gvkey", "datadate")) |>
    semi_join(llz_2018, by = "gvkey") |>
    group_by(gvkey, datadate) |>
    summarize(ratio = mean(!disclosed),
              ratio_sale = sum((!disclosed) * salecs)/sum(salecs),
              .groups = "drop") |>
    mutate(year = year(datadate)) 
```

```{r}
#| include: false
idd_periods <- get_idd_periods(min_date = "1994-01-01", max_date = "2010-12-31")
```

```{r}
#| echo: false
ciks <-
  names_seg |>
  filter(!is.na(cik)) |>
  select(gvkey, cik) |>
  collect()

state_hq_linked <- 
  state_hq |>
  inner_join(ciks, by = "cik") |>
  inner_join(disclosure, by = "gvkey", relationship = "many-to-many") |>
  filter(datadate >= min_date, datadate <= max_date) |>
  select(gvkey, datadate, ba_state) |>
  rename(state = ba_state)
```

```{r}
#| echo: false
reg_data <-
  disclosure |>
  inner_join(prin_cust_df, by = c("gvkey", "datadate")) |>
  filter(has_prin_cust) |>
  inner_join(state_hq_linked, by = c("gvkey", "datadate")) |>
  inner_join(idd_periods, 
             join_by(state, datadate >= start_date, datadate <= end_date)) |>
  filter(period_type != "Post-rejection") |>
  mutate(post = period_type == "Post-adoption",
         ln_ratio = log(1 + ratio),
         ln_ratio_sale = log(1 + ratio_sale)) |>
  select(-start_date, -end_date)
```

```{r}
#| warning: false
#| echo: false
fms <- list()
fms[[1]] <- feols(ln_ratio ~ post | gvkey + year, 
                  data = reg_data, vcov = "iid")
fms[[2]] <- feols(ln_ratio_sale ~ post | gvkey + year, 
                  data = reg_data, vcov = "iid")
```

```{r}
#| warning: false
#| echo: false
switchers <- 
  reg_data |> 
  distinct(gvkey, post) |> 
  group_by(gvkey) |> 
  filter(n() > 1) |>
  select(gvkey) |>
  distinct() |>
  ungroup()

reg_data_switchers <-
  reg_data |>
  semi_join(switchers, by = "gvkey") 

fms[[3]] <- feols(ln_ratio ~ post | gvkey + year, 
                  data = reg_data_switchers, vcov = "iid")
fms[[4]] <- feols(ln_ratio_sale ~ post | gvkey + year, 
                  data = reg_data_switchers, vcov = "iid")
```

```{r}
#| label: tbl-table-2
#| tbl-cap: Effect of IDD adoption
#| output: asis
#| echo: false
modelsummary(fms, 
             estimate = "{estimate}{stars}",
             gof_map = c("nobs", "r.squared", "adj.r.squared"),
             stars = c('*' = .1, '**' = 0.05, '***' = .01))
```

```{r}
#| echo: false
factor_t <- function(t) {
  t <- relevel(as.factor(t), ref = "-1")
  levels(t) <- gsub("^([0-9]+)", "+\\1", levels(t))
  levels(t) <- gsub("([0-9]+)", " \\1", levels(t))
  levels(t) <- gsub("^", " ", levels(t))
  t
}
```

```{r}
#| echo: false
year_diff <- function(t) {
  t <- case_when(t < -5 ~ -5,
                 t > 5 ~ 5,
                 is.na(t) ~ -Inf,
                 TRUE ~ t) 
  factor_t(t)
}
```

```{r}
#| echo: false
switch_years <-
  reg_data |>
  group_by(gvkey) |> 
  arrange(datadate) |> 
  filter(period_type == "Post-adoption" & 
           lag(period_type) == "Pre-adoption") |>
  group_by(gvkey) |>
  summarize(adoption_year = min(year), .groups = "drop")

reg_data_t <-
  reg_data |>
  left_join(switch_years, by = "gvkey") |>
  mutate(t = year_diff(year - adoption_year))
```

```{r}
#| warning: false
#| echo: false
fms <- list()
fms[[1]] <- feols(ln_ratio ~ t | gvkey + year, 
                  data = reg_data_t, vcov = "iid")
fms[[2]] <- feols(ln_ratio_sale ~ t | gvkey + year,
                  data = reg_data_t, vcov = "iid")
fms[[3]] <- feols(ln_ratio ~ t | gvkey + year,
                  data = subset(reg_data_t, !grepl("Inf", t)), vcov = "iid")
fms[[4]] <- feols(ln_ratio_sale ~ t | gvkey + year, 
                  data = subset(reg_data_t, !grepl("Inf", t)), vcov = "iid")
```

```{r}
#| label: tbl-idd-by-year
#| tbl-cap: Effect of IDD adoption by year
#| output: asis
#| echo: false
modelsummary(fms, 
             estimate = "{estimate}{stars}",
             gof_map = c("nobs", "r.squared", "adj.r.squared"),
             stars = c('*' = .1, '**' = 0.05, '***' = .01))
```

```{r}
#| echo: false
plot_coefs <- function(fm) {
  
  coefs <- 
    tibble(t = names(coef(fm)),
           value = coef(fm),
           se = se(fm)) |>
    mutate(t = as.integer(gsub("[t ]+", "", t))) |>
    filter(t != -Inf)

  ci <- 0.95
  mult <- qnorm(1 - (1 - ci)/2)

  plot <-
    tibble(t = -1, value = 0, se = 0) |>
    bind_rows(coefs) |>
    mutate(mean = value,
           top = value + mult * se,
           bot = value - mult * se) |>
    ggplot(aes(x = t, y = mean)) + 
    geom_errorbar(aes(ymin = bot, ymax = top), width = .1) +
    geom_line() +
    geom_point() +
    scale_x_continuous(breaks = seq(from = min(coefs$t), 
                                    to = max(coefs$t)))
  
  plot
}
```


```{r}
#| warning: false
#| echo: false
#| label: fig-coefs-by-year
#| fig-cap: Plot of coefficients by year
plot_coefs(fms[[1]])
```

### Discussion questions

1. The proprietary cost hypothesis tested by @Li:2018tj posits that increased cost of disclosure will lead to reduced supply of disclosure?
What rival theories exist that would make alternative predictions?
Are there other elements of disclosure theory that might be tested?
Why do you think  @Li:2018tj  focused on this specific element of the proprietary cost hypothesis?
2. In the analysis above, we do not include the control variables considered by @Li:2018tj.
For example, @Li:2018tj "include R&D expenditures to sales, advertisement [sic] expenditure to sales, and intangible assets scaled by total assets to control for a firm's proprietary costs of disclosure."
Using the approach outlined [here](https://iangow.github.io/far_book/causal-inf.html), in what circumstances would it be necessary to control for "a firm's proprietary costs of disclosure" in this way?
Do these circumstances seem applicable in this setting, where the treatment is a (presumably exogenous) shock to disclosure costs and the outcome is disclosure choices?
3. What differs between the regressions reported in columns (1) and (3), and (2) and (4) of @tbl-table-2, respectively?
What does this tell you about what drives the regression results in this setting?
What happens if you omit the year fixed effects from both sets of regressions?
What does this tell you about the role of the "non-switchers" (i.e., firms not in the `switchers` data frame) in the regression?
4. Would you expect the inclusion of controls (see the question above) to have a significant impact on the regression results?
Why or why not?
5. What differs between the regressions reported in columns (1) and (3), and (2) and (4) of @tbl-idd-by-year, respectively?
What happens if you omit the year fixed effects from both sets of regressions?
What does this tell you about what drives the regression results in this setting?
6. What patterns do you observe in the coefficients reported in @tbl-idd-by-year? 
Do these conform to what you would expect from @Li:2018tj?
(It may be easiest to focus on these in groups, e.g., those in $t-5$ to $t-2$, those in $t + 0$ to $t+3$ and those for $t+4$ and $t+5$.)
7. How do the variables in @tbl-idd-by-year differ from those used in Table 3 of @Li:2018tj?
Modify the code above (e.g., the `year_diff` function) to produce analysis closer to that reported in Table 3 of @Li:2018tj.
8. How helpful do you find @fig-coefs-by-year (plot of the coefficients by year)?
9. The `year_diff` function collapses years after $t + 5$ and before $t - 5$ into years $t + 5$ and $t - 5$, respectively. 
When would this approach make sense? 
What would be one alternative approach to handling these years?
Does your suggested approach make a difference in this setting?
10. Describe the data set created from the following code.
What proportion of the firms in the data set have `same_state` equal to `TRUE`?
For the purposes of empirical analysis, do the firms with `same_state` equal to `FALSE` enhance, or detract from, our ability to draw causal inferences about the effect of adoption of IDD?

<ul>
```{r}
switch_years <-
  reg_data_switchers |>
  group_by(gvkey) |> 
  arrange(datadate) |> 
  mutate(same_state = state == lag(state), 
         adoption_year = period_type == "Post-adoption" & 
           lag(period_type) == "Pre-adoption") |> 
  filter(adoption_year) |> 
  ungroup() 
```
</ul>

11. What issues might be implied by the following data? How might you address these?

<ul>
```{r}
reg_data_t |> count(t)
```
</ul>
